---
title: "plotting_differences_of_other_metrics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{plotting_differences_of_other_metrics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(mhwci)
devtools::load_all()
db <- mhw_connect(get_dbfile())

```

# Plotting differences of other metrics

Currently this package has functions that only compute summaries of duration, 
which 
were created at beginning as we designed the workflow. 

There is a generic function `mhw_metric_summary_sql()` 

### Example difference for one decade:

Calculate the differences of the **max of max intensity** (extremes) only for **2040 to 2049**

```{r, echo=TRUE}

# in thse two function calls, all the parameters are the same
valid_ensembles <- "006,007,008,009,010"
arise15_sql <- mhw_metric_summary_sql(mhw_table = "arise15_decade_metrics", 
                            mhw_metric = 'int_max', 
                            sql_function = 'max', 
                            start_year=2040, 
                            end_year=2049,
                            ensemble_list_string = valid_ensembles) 

arise15_raster <- summary_metrics_raster(db, arise15_sql)

## --- 
ssp245_sql <- mhw_metric_summary_sql(mhw_table = "ssp245_decade_metrics", 
                            mhw_metric = 'int_max', 
                            sql_function = 'max', 
                            start_year=2040, 
                            end_year=2049,
                            ensemble_list_string = valid_ensembles) 
  
ssp245_raster <-  summary_metrics_raster(db, ssp245_sql)
arise_max_int_diff <- ssp245_raster - arise15_raster
plot(arise_max_int_diff, main = "difference max of max intensity, \nSSP2 4.5 - ARISE 1.5, \n2040-2049")

```

### Example difference plot for 3 decades:

Calculate the differences of the **max of max intensity** 

```{r, echo=TRUE}

# we are using an 'internal' function write to be used inside an 'lapply()'  

# it's easier to set variables set above or hard-coded values in the function
# we apply, but see the `lapply()` help for how to send extra parameters

# adjust the parameters inside our function to change which metric is calculated
# takes on parameter - the one we are applying over (decade start years)
make_raster_function<- function(period_start_year ) {
      # make the sql                                
      mhw_sql <- mhw_metric_summary_sql(mhw_table = mhw_table, 
                            mhw_metric = 'int_max', 
                            sql_function = 'max', 
                            start_year = period_start_year, 
                            end_year = period_start_year + period_length_years - 1,
                            ensemble_list_string = "006,007,008,009,010") 
      
      # call the sql and create the raster
      mhw_raster <- summary_metrics_raster(db, mhw_sql)
      return(mhw_raster)
}

## --- 
# set the parameters here for the apply function
# db is already set above, otherwise use `db <- mhw_connect(dbfile)` again
start_years <- c(2040, 2050, 2060)
period_length_years  <- 10
mhw_metric  <-  'int_max' 
sql_function  <-  'max'

# set the table that the function above will use
mhw_table <-  "ssp245_decade_metrics"
ssp245_rasters <- lapply(start_years, make_raster_function)
# terra requires this step in it's own line to create a 'raster stack' of 'spat rasters'
ssp245_rasters <- terra::rast(ssp245_rasters)
# use the start years for name of list items 
names(ssp245_rasters) <- start_years

mhw_table <-  "arise15_decade_metrics"
arise15_rasters <- lapply(start_years, make_raster_function)
arise15_rasters <- terra::rast(arise15_rasters)
# use the start years for name of list items 
names(arise15_rasters) <- start_years

# I believe the lists must have the same 'names' for subtraction to work with
# a list of rasters

mhw_diffs <- ssp245_rasters - arise15_rasters
mhw_diffs


```

Plot: 

```{r echo=TRUE}
plot_rasters_squish_outliers(mhw_diffs)
```



```{r}
# cleanup connection 
DBI::dbDisconnect(db)
```
